Eye in a Disk: eyeIntegration human eye and body transcriptome database version 1.00

### 2019-01-17 manuscript.Rmd
add iovs csl

---
title: 'Eye in a Disk: eyeIntegration human eye and body transcriptome database version 1.00'
author:
  - Vinay Swamy:
      institute: ogvfb
  - David McGaughey:
      institute:
        - ogvfb
      correspondence: "yes"
      email: mcgaugheyd@mail.nih.gov
institute:
  - ogvfb: Ophthalmic Genetics and Visual Function Branch, National Eye Institute, National Institutes of Health
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: eyeIntegration_v1_app_manuscript.bib
---

@@

  From the anterior to the posterior, the human eye is composed of the cornea, lens, retina, RPE, and choroid. The development and function of these tissues is mediated through tissue and temporal specific transcript and gene expression patterns, or transcriptome. Today, RNA-sequencing (RNA-seq) is the predominant technology in quantification of the transcriptome. Analysis of gene patterns across tissue, time, and perturbation allows researchers to decipher the genetic controls of eye development and function. To this end researchers have used a wide variety of tissue types to assess gene function, including primary tissue (fetal and post-mortem), induced pluripotent stem cell (iPSC) derived cells, immortalized cell lines, and most recently, organoids. These tissue types have been deeply sequenced across the cornea [@chenIdentificationNovelMolecular2013a; @chngHighThroughputGene2013a; @chungTranscriptomicProfilingPosterior2017; @fraustoTranscriptomicAnalysisCultured2016a; @kabzaCollagenSynthesisDisruption2017; @ouyangWNT7APAX6Define2014a; @songDirectedDifferentiationHuman2016a], lens [@hanWnt5aContributesDifferentiation2018], retina [@aldiriDynamicEpigeneticLandscape2017a; @farkasTranscriptomeAnalysesHuman2013b; @hoshinoMolecularAnatomyDeveloping2017; @kaewkhawTranscriptomeDynamicsDeveloping2015; @kaewkhawTreatmentParadigmsRetinal2016; @liComprehensiveAnalysisGene2014a; @mustafiTranscriptomeAnalysisReveals2016; @pinelliAtlasGeneExpression2016; @whitmoreTranscriptomicAnalysisNasal2014b], and RPE (choroid) [@darrowDeletionDXZ4Human2016a; @harenzaTranscriptomicProfiling392017; @huIdentificationMiRNASignatures2012b; @liComprehensiveAnalysisGene2014a; @nozawaSAFARegulatesInterphase2017; @obersteinCellularResponsesHuman2017; @pengEngineeringBloodRetinalBarrier2013; @radekeRestorationMesenchymalRetinal2015a; @sainiNicotinamideAmelioratesDisease2017; @samuelAppropriatelyDifferentiatedARPE192017; @santaguidaAneuploidyinducedCellularStresses2015b; @shaoSpontaneousGenerationNovel2017; @shihRestorationMesenchymalRPE2017; @smithRetinalPigmentEpithelial2017; @stevensonGiantinknockoutModelsReveal2017; @tresiniCoreSpliceosomeTarget2015; @whewaySiRNAbasedFunctionalGenomics2015; @whitmoreTranscriptomicAnalysisNasal2014b; @auCharacterizationLincRNAExpression2017; @huIdentificationMiRNASignatures2012b]. 
  While there is a wealth of pan-eye tissue based RNA-seq data, there has not been a way for researchers to efficiently compare gene and transcript expression information across these tissues. For most human tissues the Gene Tissue Expression has generated RNA-sesq data across

>>>>

---
title: 'Eye in a Disk: eyeIntegration human eye and body transcriptome database version 1.00'
author:
  - Vinay Swamy:
      institute: ogvfb
  - David McGaughey:
      institute:
        - ogvfb
      correspondence: "yes"
      email: mcgaugheyd@mail.nih.gov
institute:
  - ogvfb: Ophthalmic Genetics and Visual Function Branch, National Eye Institute, National Institutes of Health
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: eyeIntegration_v1_app_manuscript.bib
csl: investigative-ophthalmology-and-visual-science.csl
---

@@

From the anterior to the posterior, the human eye is composed of the cornea, lens, retina, RPE, and choroid. The development and function of these tissues is mediated through tissue and temporal specific transcript and gene expression patterns, or transcriptome. Today, RNA-sequencing (RNA-seq) is the predominant technology in quantification of the transcriptome. Analysis of gene patterns across tissue, time, and perturbation allows researchers to decipher the genetic controls of eye development and function. To this end researchers have used a wide variety of tissue types to assess gene function, including primary tissue (fetal and post-mortem), induced pluripotent stem cell (iPSC) derived cells, immortalized cell lines, and most recently, organoids. These tissue types have been deeply sequenced across the cornea [@chenIdentificationNovelMolecular2013a; @chngHighThroughputGene2013a; @chungTranscriptomicProfilingPosterior2017; @fraustoTranscriptomicAnalysisCultured2016a; @kabzaCollagenSynthesisDisruption2017; @ouyangWNT7APAX6Define2014a; @songDirectedDifferentiationHuman2016a], lens [@hanWnt5aContributesDifferentiation2018], retina [@aldiriDynamicEpigeneticLandscape2017a; @farkasTranscriptomeAnalysesHuman2013b; @hoshinoMolecularAnatomyDeveloping2017; @kaewkhawTranscriptomeDynamicsDeveloping2015; @kaewkhawTreatmentParadigmsRetinal2016; @liComprehensiveAnalysisGene2014a; @mustafiTranscriptomeAnalysisReveals2016; @pinelliAtlasGeneExpression2016; @whitmoreTranscriptomicAnalysisNasal2014b], and RPE (choroid) [@darrowDeletionDXZ4Human2016a; @harenzaTranscriptomicProfiling392017; @huIdentificationMiRNASignatures2012b; @liComprehensiveAnalysisGene2014a; @nozawaSAFARegulatesInterphase2017; @obersteinCellularResponsesHuman2017; @pengEngineeringBloodRetinalBarrier2013; @radekeRestorationMesenchymalRetinal2015a; @sainiNicotinamideAmelioratesDisease2017; @samuelAppropriatelyDifferentiatedARPE192017; @santaguidaAneuploidyinducedCellularStresses2015b; @shaoSpontaneousGenerationNovel2017; @shihRestorationMesenchymalRPE2017; @smithRetinalPigmentEpithelial2017; @stevensonGiantinknockoutModelsReveal2017; @tresiniCoreSpliceosomeTarget2015; @whewaySiRNAbasedFunctionalGenomics2015; @whitmoreTranscriptomicAnalysisNasal2014b; @auCharacterizationLincRNAExpression2017; @huIdentificationMiRNASignatures2012b].

The Gene Tissue Expression (GTEx) Project has generated RNA-seq data across 53 post-mortem human tissues and 714 unique donors and presented the gene and transcript level data in a efficient web app (https://gtexportal.org/) [@carithersNovelApproachHighQuality2015; @gtexconsortiumGeneticEffectsGene2017]. Unfortunately no eye tissues are included.
